---
title: "It's getting hot in here"
subtitle: "Toronto weather data from 2013 through 2018"
author: "Anthony Ionno"
date: "June 30, 2018"
output:
  rmarkdown::html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    code_folding: hide
    self_contained: no
---
<style>
body {
    font-size: 16px;
}

</style>

## Summary 

It is hot in Toronto right now. Like 45 degrees Celsius hot. 

So, I thought I would do a short post that looks at daily weather patterns in Toronto. Both the graphs below show that since 2013 we are seeing a higher frequency of daily average temperature above zero degrees Celsius than below. 

## Analysis
### Data
The raw data used in this analysis is available on the [Environment Canada website](http://climate.weather.gc.ca/climate_data/daily_data_e.html?StationID=51459). The preprocessed version of this data is available on my [github page](https://github.com/ionnoant/2018-06-31_Toronto-weather-post). 

### R libraries
The following libraries were loaded into my R workspace for this analysis.
```{r,cache=TRUE,message=FALSE}
# Loading libraries and user defined functions
library(tidyverse);library(magrittr); library(extrafont); library(extrafontdb)
library(ggplot2); library(viridis);library(DT)

theme_ai <- function(){
  theme_minimal() +
    theme(
      text = element_text(family = "Segoe UI", color = "gray25"),
      plot.title = element_text(size=22),
      plot.subtitle = element_text(size = 20),
      axis.text = element_text(size=18, color = "gray25"),
      axis.title = element_text(size = 20),
      plot.caption = element_text(color = "gray30", size=16),
      plot.background = element_rect(fill = "gray95"),
      plot.margin = unit(c(5, 10, 5, 10), units = "mm"),
      #axis.line = element_line(color="gray50")
      axis.ticks.x = element_line(color="gray35"),
      panel.grid.major.y = element_line(colour = "gray80"),
      legend.position = "none")
  
}
```
### Preprocessing
```{r, cache=TRUE}
file_list <- list.files()
file_list <- file_list[c(5,7:11)]
df_name_list <- c("w_2018", "w_2013", "w_2014", "w_2015", "w_2016", 
                  "w_2017")

for(i in 1:length(file_list)){
  assign(df_name_list[i], read.csv(paste(file_list[i])))
}
# Had some sort of issue with the 2017 file
w_2017 <- read.csv('eng-daily-01012017-12312017.csv')
var_list <- names(w_2018)
df_list <- ls()[grep("w_",ls())]
for(i in 1:length(df_list)){
  assign(df_list[i],select(get(df_list[i]),var_list)%>%
           mutate(Date=as.Date(Date.Time, format = "%Y-%m-%d")))
}
w_2018 <- w_2018 %>%
  mutate(Date = as.Date(Date.Time, format = "%m/%d/%Y"))
w_all <- rbind(w_2013, w_2014, w_2015, w_2016, w_2017, w_2018)
w_all <- w_all[complete.cases(w_all),]
w_all <- w_all %>%
  mutate(max_temp = ifelse(abs(Min.Temp...C.) > abs(Max.Temp...C.),Min.Temp...C.,
                           Max.Temp...C.),
         temp_count = ifelse( Mean.Temp...C. > 0, 1, 0))

```

### Results 
```{r, message = FALSE, cache=TRUE, fig.width=20, fig.height=10}
# Daily Average Temp.
w_all %>%
ggplot()+
  geom_col(aes(x = Date, y = Mean.Temp...C., fill = Mean.Temp...C.<0),alpha=.7)+
  stat_smooth(aes(x = Date, y = Mean.Temp...C.),alpha = .7, se = FALSE, color = "gray55")+
  #stat_smooth(aes(x = Date, y = Max.Temp...C.))+
  #stat_smooth(aes(x = Date, y = Min.Temp...C.))+
  theme_ai()+
  scale_x_date(date_breaks = "3 month", date_labels = "%y-%m")+
  scale_y_continuous(breaks = seq(-20,40,10))+
  scale_fill_manual(values = c("seagreen3","royalblue3"))+
  labs( x = "Date (YY-MM)", y = "Average Temp. in 째C",
        title = "Daily average temperature in Toronto from 2013 through 2018",
        subtitle = "Toronto international aiport weather station used in all years",
        caption = "Data Source: Environment Canada | Creation: Anthony Ionno")

# Daily Maximum Temp.
w_all %>%
  ggplot()+
  geom_col(aes(x = Date, y = max_temp, fill = max_temp<0),alpha=.7)+
  stat_smooth(aes(x = Date, y = max_temp),alpha = .7, se = FALSE, color = "gray55")+
  #stat_smooth(aes(x = Date, y = Max.Temp...C.))+
  #stat_smooth(aes(x = Date, y = Min.Temp...C.))+
  theme_ai()+
  scale_x_date(date_breaks = "3 month", date_labels = "%y-%m")+
  scale_y_continuous(breaks = seq(-20,40,10))+
  scale_fill_manual(values = c("seagreen3","royalblue3"))+
  labs( x = "Date (YY-MM)", y = "Average Temp. in 째C",
        title = "Daily maximum temperature in Toronto from 2013 through 2018",
        subtitle = "Toronto international aiport weather station used in all years",
        caption = "Data Source: Environment Canada | Creation: Anthony Ionno")
# Table of all the data we used
datatable(select(w_all,Date, Mean.Temp...C.,max_temp)
          ,colnames = c("Date","Mean Temp. in 째C", "Max Temp. in 째C" ))
```